Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 trunk/paper_enahpc14/Sec2:COSMO-ART.tex
WS: proofread

Establishing effective energy performance benchmarking of a code under
intense development  such as \textsc{Cosmo-art} is  a challenging task
because of  the absolute necessity  that results must  be reproducible
within  an  expected  variance  for  the  duration  of  the  Exa2Green
project.  To   define  a  baseline,   it  was  necessary  to   find  a
run-configuration  capable  of   being  recreated  in  all  subsequent
versions.  Here  we give a brief  overview of the model  setup for our
performance and energy efficiency evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of  $0.22\,^{\circ}$, i.e., 50~km
in  both directions  and 40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the model initialization,  there is a period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

@@

The considered  \textsc{Cosmo-art} model  system is configured  with a
semi-Lagrangian    horizontal   advection    scheme    with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\citep{COSMO-PartI-2011}.    It  also   makes  use   of   the  Kinetic
PreProcessor solver (KPP) for  the resolution of atmospheric chemistry
ordinary  differential equations \citep{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterisations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.

>>>>

Establishing effective energy performance benchmarking of a code under
intense development  such as \textsc{Cosmo-art} is  a challenging task
because of  the absolute necessity  that results must  be reproducible
within  an  expected  variance  for  the  duration  of  the  Exa2Green
pro\-je\-ct.  To   define  a  baseline,   it  was  necessary  to   find  a
run-configuration  capable  of   being  recreated  in  all  subsequent
versions.  Here  we give a brief  overview of the model  setup for our
performance and energy efficiency evaluation.

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13$^{th}$ 2010 in order to approximately
have a half  day  of  sun  exposure and  therefore  ensure  a typical
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of  $0.22\,^{\circ}$, i.e., 50~km
in  both directions  and 40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the model initialisation there is a period of adjustment for
the model  to reach its equilibrium  state and minimise  the effect of
initial conditions for gases and aerosols on the model predictions.

@@

The considered  \textsc{Cosmo-art} model  system is configured  with a
semi-Lagrangian    horizontal   advection    sche\-me    with   tricubic
interpolation  and selective filling  diffusion option  in combination
with   the   dynamical    core   using   Runge-Kutta   time   stepping
\citep{COSMO-PartI-2011}.    It  also   makes  use   of   the  Kinetic
PreProcessor solver (KPP) for  the resolution of atmospheric chemistry
ordinary  differential equations \citep{Damian-2002}.   Concerning the
modelling  of  wet  deposition  in  aerosols, the  baseline  has  only
indirect  cloud feedbacks  but  does not  include in-cloud  scavenging
(rainout) and below-cloud  scavenging (washout) yet.  Amongst physical
parameterisations,   precipitation  formation   is   performed  by   a
two-moment cloud microphysics.
